---
title: "modifieddata"
author: "Dipseka Timsina"
date: "2024-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DirichletReg)
```

```{r setup}
updateddata <- read.csv("updateddata.csv", head=T)
pinusdata <- updateddata %>%
  select("spp", "dbh_cm", "tht_m", "stm_kg", "brk_kg", "bch_kg",
         "fol_kg", "agb_kg", "lon", "lat")

summary_table <-  pinusdata %>% group_by(spp) %>%
  summarise_all(list(min = min, max = max, mean = mean, sd=sd))
```

```{r }
PT <- pinusdata %>%
  select("spp", "dbh_cm", "tht_m", "stm_kg", "brk_kg", 
         "bch_kg", "fol_kg", "agb_kg", "lon", "lat") %>%
  filter(spp == "Pinus taeda")
```
#
# Dirichlet Regressions for estimating component biomass
# The objective is to predict the proportion of biomass in each component simultaneously
#

# Calculate proportions: 
```{r}
PT$fol_kg = PT$fol_kg/PT$agb_kg
PT$brk_kg = PT$brk_kg/PT$agb_kg
PT$bch_kg = PT$bch_kg/PT$agb_kg
PT$stm_kg = PT$stm_kg/PT$agb_kg
```

# Plot component proportions vs. DBH
```{r}
ggplot(PT) +
        geom_point(aes(x=dbh_cm, y=fol_kg, color="Foliage")) +
        geom_smooth(aes(x=dbh_cm, y=fol_kg, color="Foliage"), se=F) +
        geom_point(aes(x=dbh_cm, y=bch_kg, color="Branch")) +
        geom_smooth(aes(x=dbh_cm, y=bch_kg, color="Branch"), se=F) +
        geom_point(aes(x=dbh_cm, y=brk_kg, color="Stembark")) +
        geom_smooth(aes(x=dbh_cm, y=brk_kg, color="Stembark"), se=F) +
        geom_point(aes(x=dbh_cm, y=stm_kg, color="Stemwood")) +
        geom_smooth(aes(x=dbh_cm, y=stm_kg, color="Stemwood"), se=F) +
        theme_bw() + theme(legend.title = element_blank()) +
        xlab("Diameter at breast height (cm)") +
        ylab("Component proportions")

```
# Plot component proportions vs. tht
```{r}
ggplot(PT) +
        geom_point(aes(x=tht_m, y=fol_kg, color="Foliage")) +
        geom_smooth(aes(x=tht_m, y=fol_kg, color="Foliage"), se=F) +
        geom_point(aes(x=tht_m, y=bch_kg, color="Branch")) +
        geom_smooth(aes(x=tht_m, y=bch_kg, color="Branch"), se=F) +
        geom_point(aes(x=tht_m, y=brk_kg, color="Stembark")) +
        geom_smooth(aes(x=tht_m, y=brk_kg, color="Stembark"), se=F) +
        geom_point(aes(x=tht_m, y=stm_kg, color="Stemwood")) +
        geom_smooth(aes(x=tht_m, y=stm_kg, color="Stemwood"), se=F) +
        theme_bw() + theme(legend.title = element_blank()) +
        xlab("Total tree height (m)") +
        ylab("Component proportions")
```

# Format data for Dirichlet Regression

# Our Y values are in columns 4 - 7: 
```{r}
PT$Y = DR_data(PT[,4:7])
diri0 = DirichReg(Y ~ dbh_cm + tht_m, data=PT, model="alternative")
summary(diri0) 
```

# Add latitude and longitude in the model:
```{r}
diri1 = DirichReg(Y ~ dbh_cm + tht_m + lat + lon, data=PT, model="alternative")
summary(diri1)
```

# Compute fitted values:
```{r}
PT$pstm_kg = fitted(diri1)[,"stm_kg"]
PT$pbrk_kg = fitted(diri1)[,"brk_kg"]
PT$pbch_kg = fitted(diri1)[,"bch_kg"]
PT$pfol_kg = fitted(diri1)[,"fol_kg"]
```

# Plot Observed Vs. Predicted Component Biomass Proportions      
```{r}
op = par(oma=c(3,3,0,0), mar=c(3,3,2,2), mfrow=c(2,2))
plot(PT$stm_kg, PT$pstm_kg, font.main=1,
     main="Stem wood Biomass", pch=20); abline(0,1)
plot(PT$brk_kg, PT$pbrk_kg, font.main=1,
     main="Stem bark biomass", pch=20); abline(0,1)
plot(PT$bch_kg, PT$pbch_kg, font.main=1,
     main="Branch biomass", pch=20); abline(0,1)
plot(PT$fol_kg, PT$pfol_kg, font.main=1,
     main="Foilage biomass", pch=20); abline(0,1)
mtext(text="Observed proportion", side=1, line=0, outer=TRUE)
mtext(text="Predicted proportion", side=2, line=0, outer=TRUE)
par(op)
```

```{r}
anova(diri1, diri0)
```



# Fitted values at mean DBH, THT and LON with varying latitude
```{r}
newdat = PT
newdat$tht_m = mean(PT$tht_m)
newdat$dbh_cm = mean(PT$dbh_cm)
newdat$lon = mean(PT$lon)

newdat$stm1 = predict(diri1, newdata=newdat)[,1]
newdat$brk1 = predict(diri1, newdata=newdat)[,2]
newdat$bch1 = predict(diri1, newdata=newdat)[,3]
newdat$fol1 = predict(diri1, newdata=newdat)[,4]
```

# Plot
```{r}
ggplot(newdat) +
        geom_smooth(aes(x=lat, y=fol1, color="Foliage")) +
        geom_smooth(aes(x=lat, y=bch1, color="Branch")) +
        geom_smooth(aes(x=lat, y=brk1, color="Stembark")) +
        geom_smooth(aes(x=lat, y=stm1, color="Stemwood")) +
        theme_bw() + theme(legend.title = element_blank()) +
        xlab("Latitude") +
        ylab("Component proportions")
```

#ggsave("DiriPlot.tiff")

# Fitted values at mean DBH, THT, and LAT with varying longitude
```{r}
newdat = PT
newdat$tht_m = mean(PT$tht_m)
newdat$dbh_cm = mean(PT$dbh_cm)
newdat$lat = mean(PT$lat)

newdat$stm1 = predict(diri1, newdata=newdat)[,1]
newdat$brk1 = predict(diri1, newdata=newdat)[,2]
newdat$bch1 = predict(diri1, newdata=newdat)[,3]
newdat$fol1 = predict(diri1, newdata=newdat)[,4]
```

# Plot

```{r}
ggplot(newdat) +
        geom_smooth(aes(x=lon, y=fol1, color="Foliage")) +
        geom_smooth(aes(x=lon, y=bch1, color="Branch")) +
        geom_smooth(aes(x=lon, y=brk1, color="Stembark")) +
        geom_smooth(aes(x=lon, y=stm1, color="Stemwood")) +
        theme_bw() + theme(legend.title = element_blank()) +
        xlab("Longitude") +
        ylab("Component proportions")
```

# Diagnostic plots:
```{r}
op = par(oma=c(3,3,0,0), mar=c(3,3,2,2), mfrow=c(2,2))
plot(fitted(diri1)[,1],
     residuals(diri1, type="standardized")[,1],
     main = "Stemwood")
plot(fitted(diri1)[,2],
     residuals(diri1, type="standardized")[,2],
     main = "Stembark")
plot(fitted(diri1)[,3], 
     residuals(diri1, type="standardized")[,3],
     main = "Branch")
plot(fitted(diri1)[,4], 
     residuals(diri1, type="standardized")[,4],
     main = "Foliage")
mtext(text="Fitted values", side=1, line=0, outer=TRUE)
mtext(text="Standardized residuals", side=2, line=0, outer=TRUE)
par(op)
```

```{r calculate RMSE}
#sqrt(mean((PT$Y[,"stm_kg"]-PT$pstm_kg)^2))
#install.packages("Metrics")
#library(Metrics)
rmse_stm <- rmse(PT$Y[,"stm_kg"],PT$pstm_kg)
rmse_brk <- rmse(PT$Y[,"brk_kg"],PT$pbrk_kg)
rmse_bch <- rmse(PT$Y[,"bch_kg"],PT$pbch_kg)
rmse_fol <- rmse(PT$Y[,"fol_kg"],PT$pfol_kg)
```

```{r comparing RMSE from model0&1}
rmse_stm0 <- rmse(PT$Y[,"stm_kg"], fitted(diri0)[,"stm_kg"])
rmse_brk0 <- rmse(PT$Y[,"brk_kg"], fitted(diri0)[,"brk_kg"])
rmse_bch0 <- rmse(PT$Y[,"bch_kg"], fitted(diri0)[,"bch_kg"])
rmse_fol0 <- rmse(PT$Y[,"fol_kg"], fitted(diri0)[,"fol_kg"])

rmse_stm1 <- rmse(PT$Y[,"stm_kg"], fitted(diri1)[,"stm_kg"])
rmse_brk1 <- rmse(PT$Y[,"brk_kg"], fitted(diri1)[,"brk_kg"])
rmse_bch1 <- rmse(PT$Y[,"bch_kg"], fitted(diri1)[,"bch_kg"])
rmse_fol1 <- rmse(PT$Y[,"fol_kg"], fitted(diri1)[,"fol_kg"])
```

```{r map}

library(ggplot2)
#install.packages("maps")
library(maps)


states <- map_data("state")

ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), 
               fill = "lightgreen", color = "black") +
  geom_point(data = PT, aes(x = lon, y = lat), color = "darkgreen", size = 3) +
  labs(title = "Loblolly Pine") +
  xlab("Longitude") +
  ylab("Latitude") + 
  theme_minimal()

```

